### 2007-01-11 Documentation/part1.tex
*** empty log message ***

TOUR ITSELF (+EXAMPLE FILE) TO BE UPDATED (VALUES MAY HAVE CHANGED DUE TO AEDL FILE DEFINITION
CHANGES ETC. ALSO EXPERIMENT (OLD) NUFACT DOES NOT EXIST ANYMORE. AND: SOME ERROR IN THIS OLD
NUFACT-FILE [WW].

CHECK: GLBSETPROJECTION REALLY USED WITH GLB\_FIXED FOR DENSITY PARAMS? IS NFSTANDARD USED?

@@

One can change the minimization algorithm in \GLOBES\ with
\begin{function}
\GLBNS{SelectMinimizer}
{\tt int glbSelectMinimizer(int minimizer\_ID)} selects the minimizer {\tt GLB\_MIN\_NESTED\_POWELL}, {\tt GLB\_MIN\_DEFAULT}, or the hybrid minimzer {\tt GLB\_MIN\_POWELL} by the parameter {\tt minimizer\_ID}. 
{\tt GLB\_MIN\_NESTED\_POWELL} is the currently implemented standard algorithm, {\tt GLB\_MIN\_DEFAULT} chooses the standard algorithm at any given time, and {\tt GLB\_MIN\_POWELL} is a faster, yet rarely tested hybrid
minimizer.
\end{function}
Compared to the standard minimization which is performed for systematics first and then for the
oscillation parameters, the hybrid minimizer mixes the systematics and oscillation parameter minimizations. This method is much faster, but new (unknown) correlations between systematics parameters and oscillation parameters may emerge. Since one can change the minimizer at any time, it is recommended that one cross check the minimization for a particular experiment.

>>>>

FOR WW MYSELF: TOUR ITSELF (+EXAMPLE FILE) TO BE UPDATED (VALUES MAY HAVE CHANGED DUE TO AEDL FILE DEFINITION
CHANGES ETC. ALSO EXPERIMENT (OLD) NUFACT DOES NOT EXIST ANYMORE. AND: SOME ERROR IN THIS OLD
NUFACT-FILE [WW].

CHECK: GLBSETPROJECTION REALLY USED WITH GLB\_FIXED FOR DENSITY PARAMS? IS NFSTANDARD USED?
(QUESTIONS FOR WW MYSELF!)

@@

One can change the minimization algorithm in \GLOBES\ with
\begin{function}
\GLBNS{SelectMinimizer}
{\tt int glbSelectMinimizer(int minimizer\_ID)} selects the minimizer {\tt GLB\_MIN\_NESTED\_POWELL}, {\tt GLB\_MIN\_DEFAULT}, or the hybrid minimzer {\tt GLB\_MIN\_POWELL} by the parameter {\tt minimizer\_ID}. 
{\tt GLB\_MIN\_NESTED\_POWELL} is the currently implemented standard algorithm, {\tt GLB\_MIN\_DEFAULT} chooses the standard algorithm at any given time, and {\tt GLB\_MIN\_POWELL} is a faster, yet rarely tested hybrid
minimizer.
\end{function}
Compared to the standard minimization which is performed for systematics first and then for the
oscillation parameters, the hybrid minimizer mixes the systematics and oscillation parameter minimizations. 
This method is much faster,
but correlations between systematics parameters and oscillation parameters might lead to a different convergence behaviour in some situations. Thus, when switching to the hybrid minimizer in an application program, one has to reconsider the question whether all degeneracies are found. Since one can change the minimizer at any time, it is recommended that one cross check the minimization for a particular experiment. The danger of modifying the convergence behaviour in existing application programs is another reason why the new, faster minimizer is still declared as experimental and not used by default

### 2007-01-11 Documentation/part2.tex
*** empty log message ***

Eventually, the total event rate $x_i$ in a bin $i$ is given by
\begin{equation}
x_i(a,b,c,d)=s_i(a,b)+b_i(c,d) \, ,
\end{equation}
and is thus a function of four parameters. 
The four parameters $a,b,c,d$ have been introduced in order to describe
systematical uncertainties and are the nuisance parameters.
Each of the four parameters has a corresponding systematical error.
% The central values for all of
% the four parameters have to be always defined for built-in systematics. 
They are called  signal normalization ($a$), signal tilt/calibration ($b$), 
background  normalization ($c$) and
background tilt/calibration ($d$). Their default (central) values are zero.
% \begin{equation}
% a=1\,,\quad b=0\,,\quad c={\tt not \, \, assigned}\,,\quad d=0\,.
% \end{equation}
% Thus, for the background normalization $c$, the value has to be specified 
% in \emph{either} case. 
The errors for the normalization and the 
values of tilt/calibration are always regarded
as a pairs, \ie, they are given in the form {\tt normalization : tilt}. For example, we have\footnote{The old parameter {\tt @backgroundcenter} should not be used anymore. A background normalization center of 1.0 will be
interpreted as zero central value. IS THAT CORRECT [WW]?}
\index{aedl}{rule@{\tt rule}!signal error@{\tt "@signalerror}}
\index{aedl}{rule@{\tt rule}!background error@{\tt "@backgrounderror}}
%\index{aedl}{{\tt rule}!background center@{\tt "@backgroundcenter}}
\begin{quote}
{\tt
\tb @signalerror =       0.001  :       0.01\\
%\tb @backgroundcenter =  0.1 :       0.0\\
\tb @backgrounderror =   0.001 :       0.01
}
\end{quote}
% There is no {\tt @signalcenter} in this definition, 
% since by default the central value for the
% signal normalization is $1$ and the central value for the tilt/calibration 
% is $0$.  
%
The user has the possibility to choose the set $\{\zeta_i\}$ of nuisance 
parameters which are minimized over. This choice is specified with the 
systematics functions {\tt sys\_on\_function} and {\tt sys\_off\_function}
corresponding two systematics modes ``systematics on'' and ``systematics off''.\footnote{In earlier
versions before \GLOBES\ 3.0 the ``error dimension'' was used. The corresponding
parameters {\tt errordim\_sys\_on} and {\tt errordim\_sys\_off} are still supported
but should not be used anymore.}
\index{aedl}{rule@{\tt rule}!sys on function@{\tt "@sys\_on\_function}}
\index{aedl}{rule@{\tt rule}!sys off function@{\tt "@sys\_off\_function}}
\index{norm}{Systematics function}
The different  possibilities are shown in
\Tab~\ref{tab:error_dim}. Since the dual systematics modes define the experiment
behavior for systematics on and off, it is useful to have a matching pair of 
systematics functions for each rule 
(see also~\Sec~\ref{sec:systematics}).. The signal and background errors specified
by {\tt @signalerror} and  {\tt @backgrounderror} will then be used, if applicable.
For example, one may define 
\begin{quote}
{\tt
\tb @signalerror =       0.001  :       0.01\\
\tb @backgrounderror =   0.001 :       0.01 \\
\tb @sys\_on\_function = "chiSpectrumTilt"  \\
\tb @sys\_off\_function = "chiNoSysSpectrum"  
}
\end{quote}

For user-defined systematics \index{norm}{User-defined systematics}
(see, \Sec~\ref{sec:userchi}), one can use arbitrary names
for the systematics functions which are not pre-defined. In this case,
one would specify the systematical errors in lists, such as
\index{aedl}{rule@{\tt rule}!sys on errors@{\tt "@sys\_on\_errors}}
\index{aedl}{rule@{\tt rule}!sys off errors@{\tt "@sys\_off\_errors}}
\begin{quote}
{\tt
\tb @sys\_on\_function = "chiMySystematics"  \\
\tb @sys\_on\_errors = \{ 0.2, 0.3, 0.5 \}~// Uses three syst. errors \\
\tb @sys\_off\_function = "chiNoSysSpectrum"  \\
\tb @sys\_off\_errors = \{ \}
}
\end{quote}
WHICH SYSTEMATICAL ERRORS WILL A FUNCTION USE IF THERE IS A MIXED DECLARATION,
SUCH AS WITH SYS\_ON\_ERRORS IN SYS ON AND SIGNALERROR/BACKGROUNDERROR in SYS OFF?
DOES THAT WORK? [WW]
Note that here the systematics on and systematics off models can be used 
as fully functional different systematics modes with different systematical
errors. The interpretation of the systematical errors in {\tt @sys\_on\_errors} and
{\tt @sys\_off\_errors} is left to the application software code and the user. 
In addition, the application software has to register the user-defined systematics $\chi^2$
function with \GLB{DefineChiFunction} and uniquely identify the name
given by {\tt @sys\_on\_function} or {\tt @sys\_off\_function} to avoid
confusing different systematics routines. In addition, it is possible to define
a rule with passive systematics using {\tt glbChiZero}. Such passive systematics
is not directly used or called, but the corresponding rate vectors will be calculated
and provided for indirect access. For example, a reactor experiment with correlated
systematics between near and far detectors may define the user-defined systematics
{\tt chiReactor} for the far detector and use {\tt chiZero} for the near detector.
The routine assigned to the far detector will then perform the $\chi^2$ calculation,
which, of course, also involves the rates of the near detector. However, defining
{\tt chiReactor} in both the near and far detectors would result in a double call
of the $\chi^2$ function, \ie, the program will be a factor two slower and the
resulting $\chi^2$ will be a factor two too large.

>>>>

Eventually, the total event rate $x_i$ in a bin $i$ is given by
\begin{equation}
x_i(a,b,c,d)=s_i(a,b)+b_i(c,d) \, ,
\end{equation}
and is thus a function of four parameters. 
The four parameters $a,b,c,d$ have been introduced in order to describe
systematical uncertainties and are the nuisance parameters.
Each of the four parameters has a corresponding systematical error.
% The central values for all of
% the four parameters have to be always defined for built-in systematics. 
They are called  signal normalization ($a$), signal tilt/calibration ($b$), 
background  normalization ($c$) and
background tilt/calibration ($d$). Their default (central) values are zero.
% \begin{equation}
% a=1\,,\quad b=0\,,\quad c={\tt not \, \, assigned}\,,\quad d=0\,.
% \end{equation}
% Thus, for the background normalization $c$, the value has to be specified 
% in \emph{either} case. 
The errors for the normalization and the 
values of tilt/calibration are always regarded
as a pairs, \ie, they are given in the form {\tt normalization : tilt}. For example, we have\footnote{The old parameter {\tt @backgroundcenter} should not be used anymore. A background normalization center of 1.0 will be
interpreted as zero central value.}
\index{aedl}{rule@{\tt rule}!signal error@{\tt "@signalerror}}
\index{aedl}{rule@{\tt rule}!background error@{\tt "@backgrounderror}}
%\index{aedl}{{\tt rule}!background center@{\tt "@backgroundcenter}}
\begin{quote}
{\tt
\tb @signalerror =       0.001  :       0.01\\
%\tb @backgroundcenter =  0.1 :       0.0\\
\tb @backgrounderror =   0.001 :       0.01
}
\end{quote}
% There is no {\tt @signalcenter} in this definition, 
% since by default the central value for the
% signal normalization is $1$ and the central value for the tilt/calibration 
% is $0$.  
%
The user has the possibility to choose the set $\{\zeta_i\}$ of nuisance 
parameters which are minimized over. This choice is specified with the 
systematics functions {\tt sys\_on\_function} and {\tt sys\_off\_function}
corresponding two systematics modes ``systematics on'' and ``systematics off''.\footnote{In earlier
versions before \GLOBES\ 3.0 the ``error dimension'' was used. The corresponding
parameters {\tt errordim\_sys\_on} and {\tt errordim\_sys\_off} are still supported
but should not be used anymore.}
\index{aedl}{rule@{\tt rule}!sys on function@{\tt "@sys\_on\_function}}
\index{aedl}{rule@{\tt rule}!sys off function@{\tt "@sys\_off\_function}}
\index{norm}{Systematics function}
The different  possibilities are shown in
\Tab~\ref{tab:error_dim}. Since the dual systematics modes define the experiment
behavior for systematics on and off, it is useful to have a matching pair of 
systematics functions for each rule 
(see also~\Sec~\ref{sec:systematics}). The signal and background errors specified
by {\tt @signalerror} and  {\tt @backgrounderror} will then be used, if applicable.
For example, one may define 
\begin{quote}
{\tt
\tb @signalerror =       0.001  :       0.01\\
\tb @backgrounderror =   0.001 :       0.01 \\
\tb @sys\_on\_function = "chiSpectrumTilt"  \\
\tb @sys\_off\_function = "chiNoSysSpectrum"  
}
\end{quote}

For user-defined systematics \index{norm}{User-defined systematics}
(see, \Sec~\ref{sec:userchi}), one can use arbitrary names
for the systematics functions which are not pre-defined. In this case,
one would specify the systematical errors in lists, such as
\index{aedl}{rule@{\tt rule}!sys on errors@{\tt "@sys\_on\_errors}}
\index{aedl}{rule@{\tt rule}!sys off errors@{\tt "@sys\_off\_errors}}
\begin{quote}
{\tt
\tb @sys\_on\_function = "chiMySystematics"  \\
\tb @sys\_on\_errors = \{ 0.2, 0.3, 0.5 \}~// Uses three syst. errors \\
\tb @sys\_off\_function = "chiNoSysSpectrum"  \\
\tb @sys\_off\_errors = \{ \}
}
\end{quote}
Note that here the systematics on and systematics off models can be used 
as fully functional different systematics modes with different systematical
errors. The interpretation of the systematical errors in {\tt @sys\_on\_errors} and
{\tt @sys\_off\_errors} is left to the application software code and the user.
In addition, the application software has to register the user-defined systematics $\chi^2$
function with \GLB{DefineChiFunction} and uniquely identify the name
given by {\tt @sys\_on\_function} or {\tt @sys\_off\_function} to avoid
confusing different systematics routines. In addition, it is possible to define
a rule with passive systematics using {\tt glbChiZero}. Such passive systematics
is not directly used or called, but the corresponding rate vectors will be calculated
and provided for indirect access. For example, a reactor experiment with correlated
systematics between near and far detectors may define the user-defined systematics
{\tt chiReactor} for the far detector and use {\tt chiZero} for the near detector.
The routine assigned to the far detector will then perform the $\chi^2$ calculation,
which, of course, also involves the rates of the near detector. However, defining
{\tt chiReactor} in both the near and far detectors would result in a double call
of the $\chi^2$ function, \ie, the program will be a factor two slower and the
resulting $\chi^2$ will be a factor two too large.
Note that mixed declarations of {\tt @sys\_on\_errors}, {\tt @sys\_off\_errors}, {\tt @signalerror}, and
{\tt @backgrounderror} are possible. In this case, {\tt @sys\_on\_errors} and {\tt @sys\_off\_errors}
have priority. If for one or both of these no systematical errors are defined, {\tt @signalerror} or
{\tt @backgrounderror} will be used.
